HerMES: Spatially resolved ALMA Imaging of {\it Herschel}$^{\dagger}$-selected Dusty Star-forming Galaxies

### 2015-01-22 LaTeX/ms.tex
Duncan Farrah comments

Galaxies selected in blind surveys at far-infrared (FIR) or sub-millimieter
(sub-mm) wavelengths are generally known as dusty star-forming galaxies (DSFGs).
They are found primarily at $z \sim2$ \citep{2005ApJ...622..772C,
Bothwell:2013lr} \citep[but extend down to $z \sim 1$ and up to $z > 6$;
see][]{Messias:2014fk, Riechers:2013lr} and represent the most FIR-luminous
objects in existence during this epoch.   They are signposts of significant
over-densities \citep{Daddi:2009qy, Capak:2011qy} and likely represent the
formative stages of the most massive elliptical galaxies found in the local
Universe \citep[e.g.,][]{Ivison:2013fk, Fu:2013lr}.  Moreover, they constitute
an important component of the overall galaxy population at $z \sim 2$
\citep[e.g.,][]{2005ApJ...632..169L}, when the star-formation rate density in
the Universe peaked \citep[e.g.,][]{Lilly:1996uq, 1996MNRAS.283.1388M}.

Our collective understanding of DSFGs is currently taking a dramatic leap
forward thanks in large part to the advent of the {\it Herschel Space
Observatory} \citep[{\it Herschel};][]{Pilbratt:2010fk}.  This has resulted in a
revolution in the size and depth of blind surveys at FIR and sub-mm wavelengths.
In particular, the {\it Herschel} Multi-tiered Extragalactic Survey
\citep[HerMES;][]{Oliver:2012lr} and the {\it Herschel} Astrophysical Terahertz
Large Area Survey \citep[H-ATLAS;][]{2010PASP..122..499E} together have surveyed
$\approx 650\,$deg$^2$ at 250$\,\mu$m, 350$\,\mu$m, and 500$\,\mu$m to the
confusion limit of {\it Herschel} \citep[$\sigma \approx 6-7\,$mJy in each
band][]{Nguyen:2010fk}, plus an additional $\approx 350\,$deg$^2$ to a level
approximately double the confusion limit.  A similar effort has been undertaken
at longer wavelengths by the South Pole Telescope
\citep[SPT;][]{Carlstrom:2011qy} and the Atacama Cosmology Telescope
\citep{Swetz:2011qy}.

Theoretical expectations based on the redshift distribution and luminosity
function of DSFGs suggested that HerMES and H-ATLAS would be efficient tools for
discovering strongly lensed DSFGs \citep[e.g.,][]{1996MNRAS.283.1340B,
2007MNRAS.377.1557N}.  Submillimeter Array \citep[SMA;][]{Ho:2004lr} imaging at
870$\,\mu$m with sub-arcsecond resolution has confirmed this, with $\geq 85\%$
of the brightest sources found by {\it Herschel} ($S_{500} > 100\,$mJy) being
gravitationally lensed by an intevening galaxy or group of galaxies along the
line of sight \citep{Negrello:2010fk, Conley:2011lr, Riechers:2011uq,
Bussmann:2012lr, Wardlow:2013lr, Bussmann:2013lr}.  Sources discovered in SPT
surveys have also been shown to have a high probibility of being strongly lensed
\citep{Vieira:2013fk, Hezaveh:2013fk}.  However, statistical models
significantly over-predict the median magnification factor experienced by a {\it
Herschel} DSFG of a given $S_{500}$ \citep{Bussmann:2013lr}.  This could
indicate a deficiency in our understanding of the bright end of the intrinsic
DSFG luminosity function.

We here present Atacama Large Millimeter/submillimeter Array (ALMA) Cycle~0
imaging at 870$\,\mu$m of a sample of 29 HerMES DSFGs.
%Our targets were selected to be significantly fainter than those reported in
%\citet{Bussmann:2013lr}, so they include a larger number of unlensed or weakly
%lensed ($\mu_{870} < 2$) sources.  
Three aspects of our dataset make it unique.  First, the sample occupies a
distinct regime in flux density between the brightest {\it Herschel} DSFGs
(almost all of which are lensed) and much fainter DSFGs found in ground-based
surveys \citep[most of which are expected to be unlensed;
e.g.,][]{Hodge:2013qy}.  Second, the ALMA images are extremely sensitive
($\sigma \approx 0.2\,$mJy) and all 29 HerMES DSFGs are detected \citep[such is
not the case in previous similar studies with shallower imaging;
e.g.,][]{Smolcic:2012zl, Barger:2012yg, Hodge:2013qy}.  Third, the typical
angular resolutions are $0\farcs45$ and nearly all sources detected by ALMA are
spatially resolved.

We also obtained Gemini-South optical imaging to complement our existing array
of ancillary multi-wavelength imaging.  We use those data in this paper to
identify lensing galaxies which are typically early-types with little on-going
star-formation and therefore very weak sub-mm emission.

@@

Our sample includes 29 DSFGs drawn from five independent fields in HerMES
totaling $55\,$deg$^2$ that are accessible to ALMA and have SPIRE sensitivity
that reaches the confusion limit.  The sample is selected primarily on the
basis of $S_{500}$ and covers a range of $20 < S_{500}/{\rm (mJy)} \lesssim
100$.  Additional selection criteria based on the $S_{500}/S_{350}$ and
$S_{350}/S_{250}$ colors lead to the sample being split roughly evenly between
``blue'' (typically implying $z < 2$) and ``green'' ($z = 2-3.5$) SPIRE colors.
We caution that redshifts based strictly on SPIRE colors are uncertain due to
the degeneracy between dust temperature and redshift as well as the fact that
the multiplicity rate is high in our sample.  It is not clear that the colors
of each individual ALMA counterpart are the same as the summed colors seen by
{\it Herschel}.  Low redshift contaminants to our sample ($z < 0.1$) are
trivially removed by searching for spatially resolved counterparts in SDSS
imaging.  There is also a small contamination from blazars, which are
non-thermal emitters and are easily identified and excised from the sample
using data from the NVSS or the Very Large Array Faint Images of the Radio Sky
at Twenty-Centimeters survey \citep[FIRST;][]{Becker:1995fj}.

@@

ALMA data were obtained during Cycle~0 over a period from 2012 June to 2012
December (Program 2011.0.00539.S; PI: D. Riechers).  The observations were
carried out in good 870$\,\mu$m weather conditions which resulted in typical
system temperatures of $T_{\rm sys} \sim 130\,$K and phase fluctuations of $\sim
10\,$deg.  Each target was observed until an rms noise level of $\sigma \approx
0.2\,$mJy was achieved.  This typically required 10 minutes of on-source
integration time.  For the observations targeting the CDFS, Elais-S, and COSMOS
fields, the weather was of sufficient quality to reach $\sigma \approx
0.14\,$mJy.  The number of antennas used varied from 15 to 25.  The antennas
were configured with baseline lengths of 20$\,$m to 400$\,$m, providing a
synthesized beamsize of $\approx 0\farcs45$ FWHM while ensuring that no flux was
resolved out by the interferometer.  When possible, track-sharing of multiple
targets in a single track was used to optimize the {\it uv} coverage.

@@

We used the Sloan Digital Sky Survey (SDSS) or the 2 Micron All Sky Survey
(2MASS) to align the Gemini-S images to a common astrometric frame of
reference.  This imposes an rms uncertainty in the absolute astrometry of
$0\farcs2$ and $0\farcs4$ for SDSS and 2MASS, respectively.  The
astrometrically calibrated Gemini-S images served as the basis for aligning
higher resolution, smaller field-of-view imaging from {\it HST} or Keck that
were originally presented in \citet{Calanog:2014lr}.

@@

Sources are assumed to be elliptical Gaussians that are parameterized by the
following six free parameters: the position of the source (relative to the
primary lens if a lens is present) ($\Delta \alpha_{\rm s}$ and $\Delta
\delta_{\rm s}$), the total intrinsic flux density ($S_{\rm in}$), the
effective radius intermediate axis length ($r_{\rm s} = \sqrt{a_{\rm s} b_{\rm
s}}$), the axial ratio ($q_{\rm s}$ =  $b_{\rm s}/a_{\rm s}$), and the position
angle ($\phi_{\rm s}$, degrees east of north).  The use of an elliptical
Gaussian represents a simplification from the S\'ersic profile
\citep{1968adga.book.....S}  that is justified based on the relatively weak
constraints on the S\'ersic index found in our previous work
\citep{Bussmann:2012lr, Bussmann:2013lr}.

When an intervening galaxy (or group of galaxies) is present along the line of
sight, {\sc uvmcmcfit} accounts for the deflection of light caused by this
structure using a simple ray-tracing routine that is adopted from a 
Python routine written by A.~Bolton
\footnote{http://www.physics.utah.edu/$\sim$bolton/python\_lens\_demo/}.  This
represents a significant difference from \citet{Bussmann:2012lr} and
\citet{Bussmann:2013lr}, where we used the publicly available {\sc Gravlens}
software \citep{Keeton:2001lr} to map emission from the source plane to
the image plane for a given lensing mass distribution.  {\sc Gravlens} has a
wide range of lens mass profiles as well as a sophisticated algorithm for
mapping source-plane emission to the image-plane, but it also comes with a
significant input/output penalty that makes parallel computing prohibitively
expensive.  The use of pure-Python code for tracing the deflection of light
rays is a critical component of making {\sc uvmcmcfit} computationally
feasible.

In {\sc uvmcmcfit}, lens mass profiles are represented by $N_{\rm lens}$
singular isothermal ellipsoid (SIE) profiles, where $N_{\rm lens}$ is the
number of lensing galaxies found from the best available optical or near-IR
imaging \citep[a multitude of evidence supports the SIE as a reasonable choice;
for a recent review, see][]{Treu:2010fk}.  Each SIE is fully described by the
following five free parameters: the position of the lens relative to the
arbitrarily chosen ``image center'' based on the ALMA 870$\,\mu$m emission and
any lensing galaxies seen in the optical or near-IR ($\Delta \alpha_{\rm lens}$
and $\Delta \delta_{\rm lens}$; these can be compared with the position of the
optical or near-IR counterpart relative to the ``image center'': $\Delta
\alpha_{\rm NIR}$ and $\Delta \delta_{\rm NIR}$), the mass of the lens
(parameterized in terms of the intermediate axis angular Einstein radius,
$\theta_{\rm E}$), the axial ratio of the lens ($q_{\rm lens} = b_{\rm lens} /
a_{\rm lens}$), and the position angle of the lens ($\phi_{\rm lens}$; degrees
east of north).  Unless otherwise stated, when optical or near-IR imaging
suggests the presence of additional lenses (see Figure~\ref{fig:uvmodels}), we
estimate centroids for each lens by-eye and fix the positions of the additional
lenses with respect to the primary lens.  Each additional lens thus has 3 free
parameters: $\theta_{\rm E}$, $q_{\rm lens}$, and $\phi_{\rm lens}$.  We assume
secondary, tertiary, etc., lenses are located at the same redshift as the
primary lens.

@@

A blended source is defined as any galaxy in the mock catalogs above a
threshold flux density ($S_{\rm thresh}$) that has at least one neighbor within
a projected angular distance $d_{\rm neighbor}$.  To obtain a direct comparison
with our ALMA sample, we use $S_{\rm thresh} = 1.0\,$mJy and $d_{\rm
neighbor} = 40\arcsec$.  We use the known positions in the mock catalogs for
all blended sources and compute centroid and separations for every blended
source using the same methodology as we applied to our ALMA sample and to
ALESS.

>>>>

Galaxies selected in blind surveys at far-infrared (FIR) or sub-millimieter
(sub-mm) wavelengths are generally known as dusty star-forming galaxies
(DSFGs).  They cover a wide range in redshift from $z \sim 0.5$ to $z > 6$
\citep{2005ApJ...622..772C, Casey:2012qy, Messias:2014fk, Riechers:2013lr},
with a significant component at $z \sim 2$ \citep{Casey:2012uq,
Bothwell:2013lr}, when they represent the most FIR-luminous objects in
existence during this epoch.   They are signposts of significant over-densities
\citep{Daddi:2009qy, Capak:2011qy} and likely represent
the formative stages of the most massive elliptical galaxies found in the local
Universe \citep[e.g.,][]{Ivison:2013fk, Fu:2013lr}.  Moreover, they constitute
an important component of the overall galaxy population at $z \sim 2$
\citep[e.g.,][]{Magnelli:2011ul}, when the star-formation rate density in
the Universe peaked \citep[e.g.,][]{Lilly:1996uq, 1996MNRAS.283.1388M}.

Our collective understanding of DSFGs is currently taking a dramatic leap
forward thanks in large part to the advent of the {\it Herschel Space
Observatory} \citep[{\it Herschel};][]{Pilbratt:2010fk}.  This has resulted in
a revolution in the size and depth of blind surveys at FIR and sub-mm
wavelengths.  In particular, the {\it Herschel} Multi-tiered Extragalactic
Survey \citep[HerMES;][]{Oliver:2012lr} and the {\it Herschel} Astrophysical
Terahertz Large Area Survey \citep[H-ATLAS;][]{2010PASP..122..499E} together
have surveyed $\approx 650\,$deg$^2$ at 250$\,\mu$m, 350$\,\mu$m, and
500$\,\mu$m to the confusion limit of {\it Herschel} \citep[$\sigma \approx
6-7\,$mJy in each band][]{Nguyen:2010fk}, plus an additional $\approx
350\,$deg$^2$ to a level approximately double the confusion limit.  A similar
effort to survey large areas of the sky has been undertaken at longer
wavelengths by the South Pole Telescope \citep[SPT;][]{Carlstrom:2011qy} and
the Atacama Cosmology Telescope \citep{Swetz:2011qy}.

Theoretical expectations based on the redshift distribution and luminosity
function of DSFGs suggested that HerMES and H-ATLAS would be efficient tools for
discovering strongly lensed DSFGs \citep[e.g.,][]{1996MNRAS.283.1340B,
2007MNRAS.377.1557N}.  Submillimeter Array \citep[SMA;][]{Ho:2004lr} imaging at
870$\,\mu$m with sub-arcsecond resolution has confirmed this, with $\geq 85\%$
of the brightest sources found by {\it Herschel} ($S_{500} > 100\,$mJy) being
gravitationally lensed by an intevening galaxy or group of galaxies along the
line of sight \citep{Negrello:2010fk, Conley:2011lr, Riechers:2011uq,
Bussmann:2012lr, Wardlow:2013lr, Bussmann:2013lr}.  Sources discovered in SPT
surveys have also been shown to have a high probability of being strongly lensed
\citep{Vieira:2013fk, Hezaveh:2013fk}.  However, statistical models
significantly over-predict the median magnification factor experienced by a {\it
Herschel} DSFG of a given $S_{500}$ \citep{Bussmann:2013lr}.  This could
indicate a deficiency in our understanding of the bright end of the intrinsic
DSFG luminosity function.

We here present Atacama Large Millimeter/submillimeter Array (ALMA) Cycle~0
imaging at 870$\,\mu$m of a sample of 29 HerMES DSFGs.
%Our targets were selected to be significantly fainter than those reported in
%\citet{Bussmann:2013lr}, so they include a larger number of unlensed or weakly
%lensed ($\mu_{870} < 2$) sources.  
Three aspects of our dataset make it uniquely suited to improving our
understanding of the bright end of the intrinsic DSFG luminosity function.
First, the sample occupies a distinct regime in flux density between the
brightest {\it Herschel} DSFGs (almost all of which are lensed) and much
fainter DSFGs found in ground-based surveys \citep[most of which are expected
to be unlensed; e.g.,][]{Hodge:2013qy}.  Second, the ALMA images are extremely
sensitive ($\sigma \approx 0.2\,$mJy per beam) and all 29 HerMES DSFGs are detected
\citep[such is not the case in previous similar studies with shallower imaging;
e.g.,][]{Smolcic:2012zl, Barger:2012yg, Hodge:2013qy}.  Third, the typical
angular resolutions are $0\farcs45$ and nearly all sources detected by ALMA are
spatially resolved.

We also obtained Gemini-South optical imaging to complement our existing array
of ancillary multi-wavelength imaging.  We use those data in this paper to
identify lensing galaxies which are typically early-types with little on-going
star-formation and therefore exhibit very weak sub-mm emission.

@@

Our sample includes 29 DSFGs drawn from five independent, confusion-limited
fields in HerMES with declinations $> 2 ^\circ$ and totaling $55\,$deg$^2$.
The sample was originally selected to be a complete, flux-limited set of
objects satisfying $S_{500} > 95 \, {\rm mJy}$ that are not known radio AGN,
nearby late type galaxies, or Galactic emission.  The intention was to assemble
the largest possible sample of lenses in the HerMES fields that are accessible
to ALMA.  However, the original selection was based on the HerMES ``V5''
catalog because this was the only one available at the time of the ALMA Cycle~0
deadline.  Subsequently, improved efforts to deblend SPIRE photometry at
$500\,\mu$m were introduced that led to the development of the ``V21'' catalog
that formed the basis of the lens selection criteria used in
\citet{Wardlow:2013lr}.  A number of objects in the sample in this paper have
significantly lower $S_{500}$ values in the V21 catalog, indicating that their
$S_{500}$ in the V5 catalog was boosted by blending with nearby sources rather
than by gravitational lensing.  For this reason, the objects in this sample
comprise a combination of lenses and blends of multiple sources.

%selected primarily on the
%basis of $S_{500}$ and covers a range of $20 < S_{500}/{\rm (mJy)} \lesssim
%100$.  Additional selection criteria based on the $S_{500}/S_{350}$ and
%$S_{350}/S_{250}$ colors lead to the sample being split roughly evenly between
%``blue'' (typically implying $z < 2$) and ``green'' ($z = 2-3.5$) SPIRE colors.
%We caution that redshifts based strictly on SPIRE colors are uncertain due to
%the degeneracy between dust temperature and redshift as well as the fact that
%the multiplicity rate is high in our sample.  It is not clear that the colors
%of each individual ALMA counterpart are the same as the summed colors seen by
%{\it Herschel}.  Low redshift contaminants to our sample ($z < 0.1$) are
%trivially removed by searching for spatially resolved counterparts in SDSS
%imaging.  There is also a small contamination from blazars, which are
%non-thermal emitters and are easily identified and excised from the sample
%using data from the NVSS or the Very Large Array Faint Images of the Radio Sky
%at Twenty-Centimeters survey \citep[FIRST;][]{Becker:1995fj}.

@@

ALMA data were obtained during Cycle~0 over a period from 2012 June to 2012
December (Program 2011.0.00539.S; PI: D. Riechers).  The observations were
carried out in good 870$\,\mu$m weather conditions which resulted in typical
system temperatures of $T_{\rm sys} \sim 130\,$K and phase fluctuations of $\sim
10\,$deg.  Each target was observed until an rms noise level of $\sigma \approx
0.2\,$mJy was achieved.  This typically required 10 minutes of on-source
integration time.  For the observations targeting the CDFS, Elais-S, and COSMOS
fields, the weather was of sufficient quality to reach $\sigma \approx
0.14\,$mJy.  The number of antennas used varied from 15 to 25.  The antennas
were configured with baseline lengths of 20$\,$m to 400$\,$m, providing a
synthesized beamsize of $\approx 0\farcs45$ FWHM while ensuring that no flux was
resolved out by the interferometer (since our targets all have size scales
smaller than $1-2 \arcsec$.  When possible, track-sharing of multiple
targets in a single track was used to optimize the {\it uv} coverage.

@@

We used the Sloan Digital Sky Survey (SDSS) or the 2 Micron All Sky Survey
(2MASS) to align the Gemini-S images to a common astrometric frame of
reference.  This imposes an rms uncertainty in the absolute astrometry of
$0\farcs2$ and $0\farcs4$ for SDSS and 2MASS, respectively.  The
astrometrically calibrated Gemini-S images served as the basis for aligning
higher resolution, smaller field-of-view imaging from {\it HST} or Keck (when
available) that were originally presented in \citet{Calanog:2014lr}.

@@

Sources are assumed to be elliptical Gaussians that are parameterized by the
following six free parameters: the position of the source (relative to the
primary lens if a lens is present) ($\Delta \alpha_{\rm s}$ and $\Delta
\delta_{\rm s}$), the total intrinsic flux density ($S_{\rm in}$), the
effective radius intermediate axis length ($r_{\rm s} = \sqrt{a_{\rm s} b_{\rm
s}}$), the axial ratio ($q_{\rm s}$ =  $b_{\rm s}/a_{\rm s}$), and the position
angle ($\phi_{\rm s}$, degrees east of north).  The use of an elliptical
Gaussian represents a simplification from the S\'ersic profile
\citep{1968adga.book.....S}  that is permitted based on the relatively weak
constraints on the S\'ersic index found in our previous work
\citep{Bussmann:2012lr, Bussmann:2013lr}.

When an intervening galaxy (or group of galaxies) is present along the line of
sight, {\sc uvmcmcfit} accounts for the deflection of light caused by this
structure using a simple ray-tracing routine that is adopted from a Python
routine written by A.~Bolton
\footnote{http://www.physics.utah.edu/$\sim$bolton/python\_lens\_demo/}.  This
represents a significant difference from \citet{Bussmann:2012lr} and
\citet{Bussmann:2013lr}, where we used the publicly available {\sc Gravlens}
software \citep{Keeton:2001lr} to map emission from the source plane to the
image plane for a given lensing mass distribution.  {\sc Gravlens} has a wide
range of lens mass profiles as well as a sophisticated algorithm for mapping
source-plane emission to the image-plane, but it also comes with a significant
input/output penalty that makes parallel computing prohibitively expensive.
For example, modeling a simple system comprising one lens and one source
typically required 24-48$\,$hours using the old software, whereas the same
system can be modeled in less than one hour with the pure-Python code.  The use
of pure-Python code for tracing the deflection of light rays is a critical
component of making {\sc uvmcmcfit} computationally feasible.

In {\sc uvmcmcfit}, lens mass profiles are represented by $N_{\rm lens}$
singular isothermal ellipsoid (SIE) profiles, where $N_{\rm lens}$ is the
number of lensing galaxies found from the best available optical or near-IR
imaging \citep[a multitude of evidence supports the SIE as a reasonable choice;
for a recent review, see][]{Treu:2010fk}.  Each SIE is fully described by the
following five free parameters: the position of the lens on the sky relative to
the arbitrarily chosen ``image center'' based on the ALMA 870$\,\mu$m emission
and any lensing galaxies seen in the optical or near-IR ($\Delta \alpha_{\rm
lens}$ and $\Delta \delta_{\rm lens}$; these can be compared with the position
of the optical or near-IR counterpart relative to the ``image center'': $\Delta
\alpha_{\rm NIR}$ and $\Delta \delta_{\rm NIR}$), the mass of the lens
(parameterized in terms of the intermediate axis angular Einstein radius,
$\theta_{\rm E}$), the axial ratio of the lens ($q_{\rm lens} = b_{\rm lens} /
a_{\rm lens}$), and the position angle of the lens ($\phi_{\rm lens}$; degrees
east of north).  Unless otherwise stated, when optical or near-IR imaging
suggests the presence of additional lenses (see Figure~\ref{fig:uvmodels}), we
estimate centroids for each lens by-eye and fix the positions of the additional
lenses with respect to the primary lens.  Each additional lens thus has 3 free
parameters: $\theta_{\rm E}$, $q_{\rm lens}$, and $\phi_{\rm lens}$.  We assume
secondary, tertiary, etc., lenses are located at the same redshift as the
primary lens.

@@

A blended source is defined as any galaxy in the mock catalogs above a
threshold flux density ($S_{\rm thresh}$) that has at least one neighbor within
a projected angular distance $d_{\rm neighbor}$.  To obtain a direct comparison
with our ALMA sample, we use $S_{\rm thresh} = 1.0\,$mJy (corresponding to the
5$\sigma$ limit of the ALMA data) and $d_{\rm neighbor} = 40\arcsec$
(reflecting the size of the {\it Herschel} beam at 500$\,\mu$m).  We use the
known positions in the mock catalogs for all blended sources and compute
centroid and separations for every blended source using the same methodology as
we applied to our ALMA sample and to ALESS.
